\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2017-10-25 document/XSGRB_sample.tex
More on intro
I have written more on the introduction.

Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
derive new insights into a very wide range of topics in astrophysics. Examples
range from small scale phenomena relating to the properties of highly
relativistic jets, supernova explosions, the interstellar medium, dust extinction
curves, starburst galaxies, chemical and molecular abundances
the ionization state of the intergalactic medium, intervening
absoption systems to standard candles in cosmology.  In that sense GRB research
is very multidisciplinary. The {\it Swift} satellite \citep{Gehrels09}, which
was launched in 2004, has made it possible to harvest the great potential in
using GRBs as probes of galaxies in the intergalactic medium, which was already
hinted at by results from earlier missions \citep[e.g.][]{Ricker04}.

A crucial aspect of the great success of the {\it Swift} mission has been
the extensive ground based follow-up observations of the afterglows and of
the host galaxies of the GRBs. This work has involved a large community
of researchers both inside and outside the official {\it Swift} team. This fruitful
collaboration has been greatly facilitated by the open data access policy
of the {\it Swift} mission.

In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
was secured using low-resolution spectrographs (typically
R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. Low-resolution 
spectroscopy is powerful for a number of reasons, first of all is the ability 
to secure information for even very faint targets. This allows the 
measurements of a number of important parameters such as redshifts and spectral
slopes. For a handful of
very bright afterglows high-resolution (typically R$>$20000) spectra were
secured and for these events much more information about conditions inside the
host galaxies were extracted 
\citep[e.g.,]{}{Fiore, Thoene07, Prochaska07, Vreejswijk07}.

The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd generation
instruments at the ESO Very Large Telescope (VLT). It is a high-throughput 
echelle spectrograph
covering the full spectra range from the UV atmospheric cutoff to the near-infrared
(near-IR)
K-band with three spectral arms (UVB, VIS, NIR). It was designed very much with
transient source follow-up in mind as the fading luminosities of such sources 
makes in urgent to secure as extensive coverage as possible in the shortest possible 
time. At the same time the resolution was set to be in the range 4000--9000 in order
to be able to get a large useful spectra range between the many sky-background
emission lines in the red and near-IR.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample selection criteria and observations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Bias still, more to do, ELTs, SVOM

@@

\begin{acknowledgements}
JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
acknowledges support by the European Commission under the Marie Curie
Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
European Commission under the Marie Curie Career Integration Grant programme
(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
acknowledge expert support from the ESO staff at the Paranal and La Silla
observatories in obtaining these target of opportunity data.

>>>>

Gamma-Ray bursts (GRBs) are interesting as they provide the opportunity to
derive new insights into a very wide range of topics in astrophysics. Examples
range from small scale phenomena relating to the magnetars, properties of highly
relativistic jets, hyper/supernova explosions, the interstellar medium, dust extinction
curves, starburst galaxies, chemical and molecular abundances, escape of ionizing 
radiation, the ionization state of the intergalactic medium, intervening
absoption systems to standard candles in cosmology.  In that sense GRB research
is very multidisciplinary.

The {\it Swift} satellite \citep{Gehrels09}, which was launched in 2004, has
made it possible to harvest much of the great potential in using GRBs as probes
of galaxies in the intergalactic medium, which was already hinted at by results
from earlier missions \citep[e.g.][]{Ricker04}.  A crucial aspect of the great
success of the {\it Swift} mission has been the extensive ground based
follow-up observations of the afterglows and of the host galaxies of the GRBs.
This work has involved a large community of researchers both inside and outside
the official {\it Swift} team. This fruitful collaboration has been greatly
facilitated by the open data access policy of the {\it Swift} mission.

In the beginning of the "{\it Swift} era" most of the follow-up spectroscopy
was secured using low-resolution spectrographs (typically
R=$\lambda/\Delta\lambda$$<$1000) \citep[e.g.][]{Fynbo2009}. Low-resolution
spectroscopy is powerful for a number of reasons, first of all is the ability
to secure information for even very faint targets. This allows the measurements
of a number of important parameters such as redshifts and spectral slopes. For
a handful of very bright afterglows high-resolution (typically R$>$20000)
spectra were secured and for these events much more information about
conditions inside the host galaxies were extracted \citep[e.g.,]{}{Fiore,
Thoene07, Prochaska07, Vreejswijk07, CastroTirado2010}.

The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd
generation instruments at the ESO Very Large Telescope (VLT). It is a
high-throughput echelle spectrograph covering the full spectra range from the
UV atmospheric cutoff to the near-infrared (near-IR) K-band with three spectral
arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
in mind as the fading luminosities of such sources makes in urgent to secure as
extensive coverage as possible in the shortest possible time. At the same time
the resolution was set to be in the range 4000--9000 in order to be able to get
a large useful spectra range between the many sky-background emission lines in
the red and near-IR spectral ranges. In this paper we present the results
of a dedicated effort over the years 2009 -- 2016 to use the X-shooter
spectrograph to secure spectroscopic observations of
afterglows and host galaxies of GRBs detected by {\it Swift}. 
All data resulting from the survey are publicly availble in reduced form.

The paper is
organized in the following way: In Sect.~\ref{sample} we describe the sample
including the sample selection criteria and the observational strategy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample selection criteria and observations}\label{sample}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Bias still, more to do, ELTs, SVOM unharvested potential Kruehler? 2012, GRB08607

@@

\begin{acknowledgements}
We should add an ode to Neil Gehrels here.
JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
acknowledges support by the European Commission under the Marie Curie
Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
European Commission under the Marie Curie Career Integration Grant programme
(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
acknowledge expert support from the ESO staff at the Paranal and La Silla
observatories in obtaining these target of opportunity data.
